---
title: "20231021_Class06HW_SB"
author: "Savannah Bogus A69027475"
format: pdf
---

## Section 1:
# A
```{r}
#making a function that will apply the formula that section A uses to a vector, x
Rescale<-function(x){
  rng<-range(x)
  (x-rng[1])/(rng[2]-rng[1])
}
#testing
Rescale(1:10)
#defining the dataframe
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
```


```{r}
#applying the function to the dataframe provided
apply(df,2,Rescale)
```
# B

Can I improve this code?
```{r}
library(bio3d)
#kinase with no drug
s1<-read.pdb("4AKE")
#kinase no drug
s2<-read.pdb("1AKE")
#kinase with drug
s3<-read.pdb("1E4Y")

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

Inititally, yes, I can improve the code by getting rid of errors. 

But, I can also improve the code by writing a function that does each of the 4 parts associated with the code that was written. There's a code to read from pdb for the protein, there's a code to trim, there's a code to get just the b factor, and there's a code to plot, so my function should also have these four parts. Something that seems obvious but that I was not understanding well is that I need to define each piece as we go so that I can call it later as the function progresses to combine each part later. 

```{r}
HWfunct<-function(x){
  s_var<-read.pdb(x)
  x.chainA<-trim.pdb(s_var,chain="A",elety="CA")
  chain_bfac<-x.chainA$atom$b
  plotb3(chain_bfac,sse=x.chainA,typ="l",ylab="Bfactor")
    
}
HWfunct("1AKE")
```

```




